# Data Cleaning

```{r message=FALSE, warning=FALSE}
#load packages
library(bayesrules)
library(tidyverse)
library(janitor)
library(rstanarm)
library(bayesplot)
library(tidybayes)
library(broom.mixed)
library(modelr)
library(e1071)
library(forcats)
library(ggExtra)
library(ggpubr)
library(ggridges)
library(devtools)
library(zoo)
library(DT)
library(dplyr)
library(plyr)
Sys.setlocale("LC_TIME", "English")
```
## Load Data

```{r message=FALSE, warning=FALSE}
#load data
covid19 <- read_csv(url("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv"))
```

## Generate Monthly Incident Data

Here, we generate the daily incident data by subtracting the lagging of yesterday from the cumulative number of COVID-19 cases at present. By summing all daily incident data within one month, we generate the monthly incident data.

```{r}
covid19_filter <- covid19 %>%
  dplyr::group_by(state) %>%
  dplyr::mutate(lag1 = lag(cases, n = 1)) 

covid19_filter <- covid19_filter %>%
  mutate(lag1 = lag1 %>% 
           replace_na(0))

covid19_filter <- covid19_filter %>% mutate(case_new = cases - lag1) 
```

## Generate Time Variables

```{r message=FALSE, warning=FALSE}
covid19_month <- covid19_filter %>%
  dplyr::mutate(year = lubridate::year(date), 
                month = lubridate::month(date), 
                day = lubridate::day(date)) %>%
  dplyr::group_by(state, year, month) %>%
  dplyr::summarize(cases = sum(case_new)) 

covid19_month <- covid19_month %>%
  mutate(Date = as.yearmon(paste(covid19_month$year, covid19_month$month), "%Y %m"))
```

## Remove Incomplete Monthly Data

Since before the time we submit this project, we do not have complete data for December, 2021, we decide to remove the incomplete data from our dataset.

```{r}
covid19_month <- covid19_month %>%
  filter(!(year == 2021 & month == 12)) 
```

## Generate Daily Average Number of Cases within One Month

We take the daily average of the number of COVID-19 cases that appear within one month because different months have different days. We would like to minimize the impacts from this difference.

```{r}
covid19_month <- covid19_month %>%
  mutate(cases_daily = round(ifelse(month == 1 | month == 3 | month == 5 | month == 7 | month == 8 | month == 10 | month == 12, cases / 31, ifelse(month == 4 | month == 6 | month == 9 | month == 11, cases/ 30, ifelse(month == 2 & year == 2020, cases / 29, cases /28)))))
```

## Generate Lag in Time-Series

```{r}
covid19_month <- covid19_month %>%
  dplyr::group_by(state) %>%
  mutate(lag1 = lag(cases_daily, n = 1)) %>%
  mutate(lag3 = lag(cases_daily, n = 3)) %>%
  mutate(lag6 = lag(cases_daily, n = 6)) 

covid19_month <- covid19_month %>% 
  mutate(lag1 = lag1 %>% replace_na(0)) %>%
  mutate(lag3 = lag3 %>% replace_na(0)) %>%
  mutate(lag6 = lag6 %>% replace_na(0))
```

## Clean undefined values (0) in Logarithmic function

Since log(0) is undefined, we remove 0 from our dataset by adding one to the lagging variables.

```{r}
covid19_month <- covid19_month %>%
  mutate(lag1 = lag1 + 1) %>%
  mutate(lag3 = lag3 + 1) %>%
  mutate(lag6 = lag6 + 1)

covid19_month <- covid19_month %>%
  mutate(lag1_log = log(lag1)) %>%
  mutate(lag3_log = log(lag3)) %>%
  mutate(lag6_log = log(lag6))
```

## Generate Season Variable

```{r}
covid19_month <- covid19_month %>%
  mutate(season = ifelse(month >= 3 & month <= 5 , "spring", ifelse(month >= 6 & month <= 8, "summer", ifelse(month >= 9 & month <= 11, "fall", "winter"))))
```

## Generate Indicator Variable for Vaccination

```{r}
covid19_month <- covid19_month %>%
  mutate(vaccine = ifelse(year == 2021 & month >= 3, 1, 0))
```


## Generate Indicator Variable for Appearance of Delta Variant

```{r}
covid19_month <- covid19_month %>%
  mutate(variant = ifelse(year == 2021 & month >= 6, 1, 0))
```
